### 2020-02-12 frontmatter.tex
DLCD -> HLGD. Re-read Chp 2,3. Briefly paper 7.

The next objective of the thesis is the implementation of a minimal discrete-element lubrication/contact dynamics model for simulation of dense particle suspensions.
The main ingredients of the model include
(i) a frame-invariant, short-range lubrication model for spherical particles, and
(ii) a soft-core, stick/slide frictional contact model activated when particles overlap.
Since contact interactions dominate at high particle concentrations,
we expect the methodology to be applicable for probing the jamming of non-spherical particles and the rheology of foams as well.

@@

%
\keywords{droplets, suspension, multiphase flow, microfluidics, soft matter, rheology, depletion force, level set, ghost fluid, discrete element.}
%
\end{abstract}

@@

Nästa fokus i avhandlingen är implementeringen av en minimal diskretelementsmörjnings-/kontaktdynamikmodell för simulering av täta partikelsuspensioner. Två huvudingredienser i modellen är (i) en referensram-invariant smörjmodell med kort räckvidd för sfäriska partiklar, och (ii) en stick/slip friktionskontaktmodell med mjuk kärna som aktiveras när partiklar överlappar varandra. Eftersom kontakt-växelverkningar dominerar fysikaliskt vid höga partikelkoncentrationer, förväntar vi oss att metodologin också är tillämplig för att undersöka inklämning av icke-sfäriska partiklar och reologi för skum.

@@

\noindent
{\small (Redigerad av {\it Hanno Essén}.)}
%
\nyckelord{droppar, suspension, flerfasflöde, mikrofluidik, mjukt material, reologi, utarmningskraft, nivåuppsättning, spökvätska, diskret element.}
%
\end{abstrakt}

>>>>

The next objective of the thesis is the implementation of a minimal hybrid lubrication/granular dynamics (HLGD) model for simulation of dense particle suspensions.
The main ingredients of HLGD include
(i) a frame-invariant, short-range lubrication model for spherical particles, and
(ii) a soft-core, stick/slide frictional contact model activated when particles overlap.
Since contact interactions dominate at high particle concentrations,
we expect the methodology to be applicable for probing the jamming of non-spherical particles and the rheology of foams as well.

@@

%
\keywords{droplets, suspension, multiphase flow, microfluidics, soft matter, rheology, depletion force, level set, ghost fluid.}
%
\end{abstract}

@@

Nästa fokus i avhandlingen är implementeringen av en minimal hybrid lubrication/granular dynamics (HLGD) för simulering av täta partikelsuspensioner. Två huvudingredienser i modellen är (i) en referensram-invariant smörjmodell med kort räckvidd för sfäriska partiklar, och (ii) en stick/slip friktionskontaktmodell med mjuk kärna som aktiveras när partiklar överlappar varandra. Eftersom kontakt-växelverkningar dominerar fysikaliskt vid höga partikelkoncentrationer, förväntar vi oss att metodologin också är tillämplig för att undersöka inklämning av icke-sfäriska partiklar och reologi för skum.

@@

\noindent
{\small (Redigerad av {\it Hanno Essén}.)}
%
\nyckelord{droppar, suspension, flerfasflöde, mikrofluidik, mjukt material, reologi, utarmningskraft, nivåuppsättning, spökvätska.}
%
\end{abstrakt}

### 2020-02-12 overview.tex
DLCD -> HLGD. Re-read Chp 2,3. Briefly paper 7.

\thesisstructure
In the remaining of the thesis, we summarise previous studies in the literature to provide an extended overview on droplet interactions and suspension flow.
This is the \emph{Part I} of the thesis and contains four more chapters.
In Chapter \ref{chp:micro-hydro}, we formulate the equations governing the microhydrodynamics and
spotlight the essential physics for one, two and more particles in various flow fields or geometries.
They are the prerequisites for understanding droplet interactions in more complicated conditions.
In Chapter \ref{chp:num-methods}, we introduce two categories of numerical methodologies that
either solve the governing equations directly or approximate the solution in reduced order.
Specifically, each category is divided into several classes, which have further variations and are mentioned succinctly.
These are followed by a short detour in Chapter \ref{chp:misc},
where two additional topics studied alongside the main subject are motivated and described.
Finally, Chapter \ref{chp:summary} concludes the first part of the thesis,
leading us to \emph{Part II}, where the original results are collected in the form of papers.

@@

This chapter summarises the mathematical formulation and theoretical basis of microhydrodynamics and particle interactions.
By \emph{particle}, we loosely mean a solid or liquid body of micron to millimetre in size, which may also be referred to as \emph{droplet} for abuse of language.
As preliminaries, most of the contents below are given without any proof. The interested reader is encouraged to read \cite{Batchelor, hb, ps, kim_karrila, graham_2018} for greater details.

@@

When the number of particles is large within a suspension, analytical calculations of their relative motions can quickly become forbidden due to the complexities of the $N$-body problem.\footnote{Theoretically, $N = 3$ is the lower bound to break the time reversal symmetry.}
This is because, despite the fact that we know how to calculate the interaction at the individual level, the collective dynamics are often in regimes that are far from equilibrium, for which there is no established theoretical framework to rely on.
Even if we are primarily interested in averaged properties of a suspension, \emph{a priori} predictions can seldom be made as the \emph{microstructure}, \ie the spatial arrangement of each suspending particle, both determines and is determined by the bulk motion in a dynamic way \citep{Brady_Bossis1988}.
Nevertheless, in the following, we will introduce a few macroscopic properties of particle suspensions and show how they can be related to certain measurable or computable system parameters. The purpose is to provide a basic physical understanding and, in particular, the rheological characterizations of a particulate system.

@@

The $N$-particle diffusion tensor $\bm D$ can be generalized from the Einstein relation \citep{Einstein1905} and is given by
\begin{equation}
 \begin{aligned}
   {\bm D} = k_BT{\bm M},
 \end{aligned}
\end{equation}
where $k_B$ is the Boltzmann constant, $T$ is the temperature, and ${\bm M}$ is again the mobility tensor. That is, diffusion and mobility have the same qualitative meaning for a particulate system.

For non-Brownian suspensions, several particle diffusivities may be defined, though the motion of a particle is typically not diffusive except at short and long times \citep{Brady_Bossis1988}.
In the short-time limit, the self-diffusivity ${\bm D}_0^s$ can be taken as the diagonal of $\bm{D}$ per particle and average over all particles; whereas in the long-time limit, the self-diffusivity ${\bm D}_\infty^s$ is usually defined with the mean-square displacement, \viz
\begin{equation}
 \begin{aligned}
   {\bm D}_\infty^s = \lim_{t \to \infty} \frac{ \langle \big({\bm x}_i - {\bm x}_{i,0} )^2 \rangle_i }{2t},
 \end{aligned}
\end{equation}
where ${\bm x}_i$ is the position of particle $i$ at $t \to \infty$, ${\bm x}_{i,0}$ is its starting position, and $\langle \cdot \rangle_i$ is an average over $i$. Note that self-diffusivities are vectors of the same size of the dimensionality of the space, $d$. For an isotropic suspension, it can also be expressed as a scalar by averaging the components of ${\bm D}^s$ over $d$.

@@

One of the central tasks of rheology is to determine the bulk stress tensor $\langle \bm{\Sigma} \rangle_i$ of a flowing material, expressed as \citep{batchelor_1970, batchelor_1977}
\begin{equation} \label{eq:stress-tensor}
  \begin{aligned}
    \langle \bm{\Sigma} \rangle_i = -\bm{\Pi} +2\mu \bm{E}^\infty + \frac{N}{V} 
    \bigg(\langle \bm{S}^H \rangle_i + \langle \bm{S}^B \rangle_i +\langle \bm{S}^A \rangle_i \bigg), 
  \end{aligned}
\end{equation} 
where $V$ is the enclosing volume,
$N$ is the number of particles,
and $\bm{S}$'s denote the stresslets due to various particle interactions, including the hydrodynamic ($H$), Brownian ($B$), and any other applied ($A$) forces, as written.
Apart from the viscous stress tensor, $-\bm{\Pi} +2\mu \bm{E}^\infty$, due to the underlying flow, the rest of the terms in Eq.\ \eqref{eq:stress-tensor} are the particle contribution.
In general, these stress components are all related to the suspension microstructure.
In the case of the hydrodynamic stresslet, it can be computed as
\begin{equation} 
  \begin{aligned}
    \langle \bm{S}^H \rangle_i = - \langle \bm{R}_{SU}\cdot(\bm{U}-\bm{U}^\infty) - \bm{R}_{SE}:\bm{E}^\infty \rangle_i , 
  \end{aligned}
\end{equation}
where $\bm{R}_{SU}$ and $\bm{R}_{SE}$ are configuration-dependent resistance matrices that resemble $\bm{M}^{-1}$.
See Section \ref{subsec:sd} and \emph{Paper 7} for further details.

@@

Instead of meshing and tracking the moving boundary on a separate grid, an alternative approach is to capture the particle motion on the same grid at the expanse of an additional function. Such function may correspond to the local volume fraction of a dispersed phase, or measure a geometric quantity of the interface away from the particle.
Conceptually, it specifies the multiphase flow in a full Eulerian manner, and is sometimes referred to as \emph{Eulerian-Eulerian} methods (front-tracking methods would be called \emph{Eulerian-Lagrangian} methods).
Many numerical methods fall under the Eulerian-Eulerian framework. In the following, we describe three of them (see Figure \ref{fig:front-capture}).
As always, a brief historical background is given first, followed by the definition of the ``front-capturing'' function and its governing equation.
We also highlight the major advantages and disadvantages of each method on a general basis.
Finally, when coupling these front-capturing methods with a flow solver, the internal boundary conditions need to be imposed near the interface and, in principle, can be treated in exactly the same way as in front-tracking methods, cf.\ Eq.\ \eqref{eq:surf-tension}. We skip these details and refer the reader to \cite{Brackbill_JCP_1992, Fedkiw_JCP_1999, Lalanne_JCP_2015, Popinet_ARFM_2018, Gibou_Fedkiv_Osher, ICLS} for further discussions.

@@

Another way to model interface evolution is to use level set (LS) methods \citep{sethian_1999a}.
Originally devised by \cite{Osher_Sethian_levelset}, LS methods encompass a broad class of algorithms and numerical schemes for ``front propagation'' with curvature-dependent speeds.
The mathematical machinery is rooted in the boundary-value and initial-value partial-differential-equations (PDE) perspective on moving interfaces.\footnote{It may seem similar to VOF methods in the sense that the both methods are Eulerian and the governing equations can be written in a similar way. However, the level set method embeds the \emph{geometry} of an interface, while VOF has the \emph{volume} embedding. This, together with that level set is Lipschitz-continuous and VOF is discontinuous, makes the two approaches fundamentally different. See Figure \ref{fig:front-capture}.}
Its applications include geometry, computer vision, fluid mechanics, combustion, tissue growth, path optimization, \etc
In terms of fluid mechanics, specifically, the last three decades has witnessed continuous numerical development in applying level set methods to multiphase flow, see \eg \cite{Mulder_JCP_1992, Sussman_JCP_1994, Sussman_JCP_2000, Enright_JCP_2002, Olsson_JCP_2005, Marchandise_JCP_2007, Desjardins_JCP_2008, Desjardins_JCP_2009, Aanjaneya_JCP_2013, Luo_JCP_2015, ICLS}. See also \cite{Gibou_Fedkiv_Osher} for a recent review.

@@

Geometrically, the normal and the curvature of the interface can be readily derived from the level set as below,\footnote{Some authors express the curvature as $\kappa = -\nabla \cdot {\bm n}$. This is merely due to different sign conventions for the level set.}
\begin{subequations} \label{eq:ls-geo}
 \begin{equation}
    {\bm n} = \frac{\nabla \phi}{\abs{ \nabla \phi }},
 \end{equation}
 \begin{equation} \label{eq:curv}
    \kappa = \nabla \cdot {\bm n}.
 \end{equation}
\end{subequations}

When solving the level set equation for multiphase flow, the most natural choice for $F$ is the local flow velocity projected in the normal direction, \ie $F=\bm{u \cdot n}$, reducing Eq.\ \eqref{eq:ls-eq} to
\begin{equation} \label{eq:ls-avd}
  \frac{\partial \phi}{\partial t} + \bm{u} \cdot \nabla \phi = 0.
\end{equation}
However, any non-uniform flow will distort the LS field, making Eq.\ \eqref{eq:ls-def} invalid after a few advection steps and making the numerical solution unstable.
To remedy this problem, it is customary to perform an infrequent level set ``reinitialization'' to reshape $\phi$ into a distance function while preserving the interface location \citep{Sussman_JCP_1994}.
Various techniques have been developed to do so by \eg using fast marching methods \citep{sethian_1999a}, or solving a time-dependent Hamilton-Jacobi equation \citep{Sussman_JSC_1997, Russo_JCP_2000}. It is also possible to avoid this additional step by modifying $F$ such that the gradient of $\phi$ is always fixed, see \eg \cite{Adalsteinsson_JCP_1999}.

At this point, it is important to mention a critical numerical issue of classical level set methods. The problem is, even if Eq.\ \eqref{eq:ls-avd} is solved on an incompressible velocity field, there is no guarantee that the numerical solution will conserve the total enclosed volume; in practice, there is often significant ``mass loss''.
The reason for this problem is a combination of numerical dissipation, under-resolution, and lack of built-in mass conservation properties in the numerical scheme.
Plenty of methods have thus been proposed to overcome such issues by using higher-order numerical schemes, coupling the LS with VOF or lagrangian particles, imposing additional mass constraints, \etc See \cite{ICLS} (\emph{Paper 1}) for a brief overview.

@@

Finally, we summarise the numerical algorithm of the interface-correction level set/ghost fluid method of \emph{Paper 1} in Algorithm \ref{al:ls}.
Here, the level set field is integrated in time with a total-variation-diminishing third-order Runge-Kutta scheme; the spatial derivatives, ${AD}$, are computed with either fifth-order upstream-central (HOUC) or weighted essentially non-oscillatory (WENO) schemes. The mass correction and reinitialization are treated similarly.
In the flow solver, the pressure Poisson equation is split into a constant-coefficient part and a variable-coefficient part to utilise the fast Fourier transform.
The surface tension is incorporated in the same equation using the ghost fluid method (denoted by $[\cdot]_\Gamma$).
The full algorithm is presented in details in \emph{Paper 1}. Note the high-level simplicity of the level set method comparing to the immersed boundary method (Algorithm \ref{al:ibm}).

@@

The last class of methods that we will describe is derived from theories of capillarity and critical phenomena.
Here, by introducing a so-called order parameter, a single-component fluid with varying density or a binary fluid with two immiscible components are modelled by a phase field separated by an interface of finite thickness \citep{Anderson_McFadden_Wheeler}.
Historically, the proposition of a finite-width interface dates back to as early as \cite{Poisson1831}. Since then, various theories have been developed by Maxwell, Gibbs, Lord Rayleigh, van der Waals, etc.; see \eg \cite{van-der-waals1893, Cahn1961}.
We note that the physical notion of finite width is particularly relevant for fluids that are near their critical point, or for processes involving moving contact lines, breakup or coalescence, where a sharp interface description will be faced with singularities \citep{Zhang_Mohseni_2018, Eggers1997}.
For this reason, phase field methods are also regarded as \emph{diffuse interface} methods.

@@

In particle-based methods, the central equation to solve is
\begin{equation} 
 \begin{aligned} \label{eq:force-balance}
  {\bm m} \cdot \frac{d{\bm U}}{dt} = {\bm F}^H + {\bm F}^A, 
 \end{aligned}
\end{equation}
where ${\bm m}$ is a generalized mass/moment-of-inertia matrix,
${\bm U}$ is the particle translational/rotational velocity vector,
and ${\bm F}$ represents the force/torque vectors owing to hydrodynamics (denoted with superscript $^H$)
or any applied interactions (denoted with superscript $^A$) that may include contact, electrostatic, van der Waals, or magnetic forces, etc.%
\footnote{In principle, stochastic forces can also be included in the general formulation,
though its numerical treatments demand special care such that the fluctuation-dissipation theorem is satisfied.}
Note that, for $N$ particles in 3D, both ${\bm U}$ and ${\bm F}$ have dimension $6N$, while ${\bm m}$ has size $6N \times 6N$.

@@

There is one more crucial step in the Stokesian dynamics method.
In the construction of $\mathscr{M}^\infty$, lubrication interactions are not fully accounted for due to the omission of higher-order multipole moments (hence the superscript $^\infty$). To correctly include all contributions from the lubrication, $\mathscr{R}$ is calculated as
\begin{equation} \label{eq:sd-lub-correction}
 \begin{aligned}
  \mathscr{R} = (\mathscr{M}^\infty)^{-1} +\mathscr{R}_{2B} - \mathscr{R}_{2B}^\infty,
 \end{aligned}
\end{equation}
where $\mathscr{R}_{2B}^\infty$ denotes the far-field parts of lubrication already embedded in $\mathscr{M}^\infty$, and the subscript $_{2B}$ indicates that lubrication is mainly a two-body interaction. See \cite{durlofsky_brady_bossis_1987, Brady_Bossis1988} for more detailed discussions.

Finally, the solution procedure of the Stokesian dynamics method can be summarised as follows.
\begin{enumerate}
\item Build the grand mobility matrix $\mathscr{M}^\infty$ from the suspension microstructure.
\item Obtain the grand resistance matrix $\mathscr{R}$ by inverting $\mathscr{M}^\infty$ and adding the lubrication correction as in Eq.\ \eqref{eq:sd-lub-correction}.
\item Compute the particle velocity according to Eq.\ \eqref{eq:sd-vel}.
\item Update the particle position using a temporal integration scheme, such as the Euler's method $\bm{x}(t+\Delta t)= \bm{x}(t) +\bm{U} \Delta t$.
\item Go back to Step 1 to continue the time integration unless the desired time step has been reached.
\end{enumerate}

@@

The above expressions for the contact forces and torques can be directly added in the ${\bm F}^A$ vector in Eq.\ \eqref{eq:force-balance}
and be solved similarly to the Stokesian dynamics method.
From a computational standpoint, the present method does not require to invert the mobility matrix,
and the grand resistance matrix composed of pair interactions is sparse;
hence, it can be solve much faster numerically.
Alternatively, the particle dynamics can also be obtained by explicitly solving Eq.\ \eqref{eq:force-balance}.
That is, first computing all forces and torques to get the acceleration, then integrating in time once to obtain the velocity and twice to update the position.
Similar techniques have been called dissipative particle dynamics or Brownian dynamics, depending on the context \citep{Hoogerbrugge_1992, Groot_Warren_1997}.
Provided that the particle Reynolds number remains much less than unity, this is equivalent to the quasi-static matrix inversion approach used in Stokesian dynamics.

@@

In this chapter, we describe two additional topics that were investigated during the course of the present work.
As side projects, they share with the rest of the thesis in that the flow involved are multiphase and characterised by relatively small length scales; hence, they can be, and have been, studied using similar numerical methods.
Beyond that, there is no further connection.
In the following, we first give the physical motivation of each problem, then summarise the overall understanding and research methodology prior to our study.
The new results are reported in \emph{Paper 5 \& 6}.

@@

Finally, as an outlook, we summarise the lessons learned during all the projects and name a few open questions worth further investigation:
\medskip

>>>>

\thesisstructure
In the remaining of the thesis, we summarize previous studies in the literature to provide an extended overview on droplet interactions and suspension flow.
This is the \emph{Part I} of the thesis and contains four more chapters.
In Chapter \ref{chp:micro-hydro}, we formulate the equations governing the microhydrodynamics and
spotlight the essential physics for one, two and more particles in various flow fields or geometries.
They are the prerequisites for understanding droplet interactions in more complicated conditions.
In Chapter \ref{chp:num-methods}, we introduce two categories of numerical methodologies that
either solve the governing equations directly or approximate the solution in reduced order.
Specifically, each category is divided into several classes, which have further variations and are mentioned succinctly.
These are followed by a short detour in Chapter \ref{chp:misc},
where two additional topics studied alongside the main subject are motivated and described.
Finally, Chapter \ref{chp:summary} concludes the first part of the thesis,
leading us to \emph{Part II}, where the original results are collected in the form of papers.

@@

This chapter summarizes the mathematical formulation and theoretical basis of microhydrodynamics and particle interactions.
By \emph{particle}, we loosely mean a solid or liquid body of micron to millimetre in size, which may also be referred to as \emph{droplet} for abuse of language.
As preliminaries, most of the contents below are given without any proof. The interested reader is encouraged to read \cite{Batchelor, hb, ps, kim_karrila, graham_2018} for greater details.

@@

When the number of particles is large within a suspension, analytical calculations of their relative motions can quickly become forbidden due to the complexities of the $N$-body problem.\footnote{Theoretically, $N = 3$ is the lower bound to break the time reversal symmetry.}
This is because, despite the fact that we know how to calculate the interaction at the individual level, the collective dynamics are often in regimes that are far from equilibrium, for which there is no established theoretical framework to rely on.
Even if we are primarily interested in averaged properties of a suspension, \emph{a priori} predictions can seldom be made as the \emph{microstructure}, \ie the spatial arrangement of each suspending particle, both determines and is determined by the bulk motion in a dynamic way \citep{Brady_Bossis1988}.
Nevertheless, in the following, we will introduce a few macroscopic properties of particle suspensions and show how they can be related to certain measurable or computable system parameters. The purpose is to provide a basic physical understanding and, in particular, the rheological characterisations of a particulate system.

@@

The $N$-particle diffusion tensor $\bm D$ can be generalised from the Einstein relation \citep{Einstein1905} and is given by
\begin{equation}
 \begin{aligned}
   {\bm D} = k_BT{\bm M},
 \end{aligned}
\end{equation}
where $k_B$ is the Boltzmann constant, $T$ is the temperature, and ${\bm M}$ is again the mobility tensor. That is, diffusion and mobility have the same qualitative meaning for a particulate system.

For non-Brownian suspensions, several particle diffusivities may be defined, though the motion of a particle is typically not diffusive except at short and long times \citep{Brady_Bossis1988}.
In the short-time limit, the self-diffusivity ${\bm D}_0^s$ can be taken as the diagonal of $\bm{D}$ per particle and average over all particles; whereas in the long-time limit, the self-diffusivity ${\bm D}_\infty^s$ is usually defined with the mean-square displacement, \viz
\begin{equation}
 \begin{aligned}
   {\bm D}_\infty^s = \lim_{t \to \infty} \frac{ \big \langle \big({\bm x}_i - {\bm x}_{i,0} )^2 \big \rangle_i }{2t},
 \end{aligned}
\end{equation}
where ${\bm x}_i$ is the position of particle $i$ at $t \to \infty$, ${\bm x}_{i,0}$ is its starting position, and $\langle \cdot \rangle_i$ is an average over $i$. Note that self-diffusivities are vectors of the same size of the dimensionality of the space, $d$. For an isotropic suspension, it can also be expressed as a scalar by averaging the components of ${\bm D}^s$ over $d$.

@@

One of the central tasks of rheology is to determine the bulk stress tensor $\langle \bm{\Sigma} \rangle_i$ of a flowing material, expressed as \citep{batchelor_1970, batchelor_1977}
\begin{equation} \label{eq:stress-tensor}
  \begin{aligned}
    \langle \bm{\Sigma} \rangle_i = -\bm{\Pi} +2\mu \bm{E}^\infty + \frac{N}{V} 
    \bigg( \big \langle \bm{S}^H  \big \rangle_i +  \big \langle \bm{S}^B  \big \rangle_i + \big \langle \bm{S}^A  \big \rangle_i \bigg), 
  \end{aligned}
\end{equation} 
where $V$ is the enclosing volume,
$N$ is the number of particles,
and $\bm{S}$'s denote the stresslets due to various particle interactions, including the hydrodynamic ($H$), Brownian ($B$), and any other applied ($A$) forces, as written.
Apart from the viscous stress tensor, $-\bm{\Pi} +2\mu \bm{E}^\infty$, due to the underlying flow, the rest of the terms in Eq.\ \eqref{eq:stress-tensor} are the particle contribution.
In general, these stress components are all related to the suspension microstructure.
In the case of the hydrodynamic stresslet, it can be computed as
\begin{equation} 
  \begin{aligned}
    \langle \bm{S}^H \rangle_i = - \langle \bm{R}_{SU}\cdot(\bm{U}-\bm{U}^\infty) - \bm{R}_{SE}:\bm{E}^\infty \rangle_i , 
  \end{aligned}
\end{equation}
where $\bm{R}_{SU}$ and $\bm{R}_{SE}$ are configuration-dependent resistance matrices that resemble $\bm{M}^{-1}$.
See Section \ref{subsec:sd} and \emph{Paper 7} for further details.

@@

Instead of meshing and tracking the moving boundary on a separate grid,
an alternative approach is to capture the particle motion on the same grid at the expanse of an additional function.
Such function may correspond to the local volume fraction of a dispersed phase, or measure a geometric quantity of the interface away from the particle.
Conceptually, it specifies the multiphase flow in a full Eulerian manner, and is sometimes referred to as \emph{Eulerian-Eulerian} methods (front-tracking methods would be called \emph{Eulerian-Lagrangian} methods).

Many numerical methods fall under the Eulerian-Eulerian framework.
In the following, we describe three of them (see Figure \ref{fig:front-capture}).
As always, a brief historical background is given first, followed by the definition of the ``front-capturing'' function and its governing equation.
We also highlight the major advantages and disadvantages of each method on a general basis.
Finally, when coupling these front-capturing methods with a flow solver,
the internal boundary conditions need to be imposed near the interface and, in principle,
can be treated in exactly the same way as in front-tracking methods, cf.\ Eq.\ \eqref{eq:surf-tension}.
We skip these details and refer the reader to \cite{Brackbill_JCP_1992, Fedkiw_JCP_1999, Lalanne_JCP_2015, Popinet_ARFM_2018, Gibou_Fedkiv_Osher, ICLS} for further discussions.

@@

Another way to model interface evolution is to use level set (LS) methods \citep{sethian_1999a}.
Originally devised by \cite{Osher_Sethian_levelset}, LS methods encompass a broad class of algorithms and numerical schemes for ``front propagation'' with curvature-dependent speeds.
The mathematical machinery is rooted in the boundary-value and initial-value partial-differential-equations (PDE) perspectives on moving interfaces.%
\footnote{LS may seem similar to VOF methods in the sense that both methods are Eulerian and the governing equations can be written in a similar way. However, the level set embeds the \emph{geometry} of an interface, while VOF has the \emph{volume} embedding. This, together with that level set is Lipschitz-continuous and VOF is discontinuous, makes the two approaches fundamentally different. See Figure \ref{fig:front-capture}.}
Its applications include geometry, computer vision, fluid mechanics, combustion, tissue growth, path optimization, \etc
In terms of fluid mechanics, specifically, the last three decades has witnessed continuous numerical development in applying level set methods to multiphase flow,
see \eg \cite{Mulder_JCP_1992, Sussman_JCP_1994, Sussman_JCP_2000, Enright_JCP_2002, Olsson_JCP_2005, Marchandise_JCP_2007, Desjardins_JCP_2008, Desjardins_JCP_2009, Aanjaneya_JCP_2013, Luo_JCP_2015, ICLS}.
See also \cite{Gibou_Fedkiv_Osher} for a recent review.

@@

Geometrically, the normal and the curvature of the interface can be readily derived from the level set as below,%
\footnote{Some authors express the curvature as $\kappa = -\nabla \cdot {\bm n}$. This is merely due to different sign conventions for the level set.}
\begin{subequations} \label{eq:ls-geo}
 \begin{equation}
    {\bm n} = \frac{\nabla \phi}{\abs{ \nabla \phi }},
 \end{equation}
 \begin{equation} \label{eq:curv}
    \kappa = \nabla \cdot {\bm n}.
 \end{equation}
\end{subequations}

When solving the level set equation for multiphase flow, the most natural choice for $F$ is the local flow velocity projected in the normal direction,
\ie $F=\bm{u \cdot n}$, reducing Eq.\ \eqref{eq:ls-eq} to
\begin{equation} \label{eq:ls-avd}
  \frac{\partial \phi}{\partial t} + \bm{u} \cdot \nabla \phi = 0.
\end{equation}
However, any non-uniform flow will distort the LS field, making Eq.\ \eqref{eq:ls-def} invalid after a few advection steps and making the numerical solution unstable.
To remedy this problem, it is customary to perform an infrequent level set ``reinitialization'' to reshape $\phi$ into a distance function while preserving the interface location \citep{Sussman_JCP_1994}.
Various techniques have been developed to do so by using fast marching methods \citep{sethian_1999a} or solving a time-dependent Hamilton-Jacobi equation \citep{Sussman_JSC_1997, Russo_JCP_2000}.
It is also possible to avoid this additional step by modifying $F$ such that the gradient of $\phi$ is always fixed, see \eg \cite{Adalsteinsson_JCP_1999}.

At this point, it is important to mention a critical numerical issue of classical level set methods.
The problem is, even if Eq.\ \eqref{eq:ls-avd} is solved on an incompressible velocity field,
there is no guarantee that the numerical solution will conserve the total enclosed volume;
in practice, there is often significant ``mass loss''.
The reason for this problem is a combination of numerical dissipation, under-resolution and lack of built-in mass conservation properties in the numerical scheme.
Plenty of methods have thus been proposed to overcome such issues by using higher-order numerical schemes, coupling the LS with VOF or lagrangian particles, imposing additional mass constraints, \etc See \cite{ICLS} (\emph{Paper 1}) for a brief overview.

@@

Finally, we summarize the numerical algorithm of the interface-correction level set/ghost fluid method of \emph{Paper 1} in Algorithm \ref{al:ls}.
Here, the level set field is integrated in time with a total-variation-diminishing third-order Runge-Kutta scheme; the spatial derivatives, ${AD}$, are computed with either fifth-order upstream-central (HOUC) or weighted essentially non-oscillatory (WENO) schemes. The mass correction and reinitialization are treated similarly.
In the flow solver, the pressure Poisson equation is split into a constant-coefficient part and a variable-coefficient part to utilise the fast Fourier transform.
The surface tension is incorporated in the same equation using the ghost fluid method (denoted by $[\cdot]_\Gamma$).
The full algorithm is presented in details in \emph{Paper 1}. Note the high-level simplicity of the level set method comparing to the immersed boundary method (Algorithm \ref{al:ibm}).

@@

The last class of methods that we will describe is derived from theories of capillarity and critical phenomena.
Here, by introducing a so-called order parameter, a single-component fluid with varying density or a binary fluid with two immiscible components are modelled by a phase field separated by an interface of finite thickness \citep{Anderson_McFadden_Wheeler}.
Historically, the proposition of a finite-width interface dates back to as early as \cite{Poisson1831}. Since then, various theories have been developed by Maxwell, Gibbs, Lord Rayleigh, van der Waals, etc.; see \eg \cite{van-der-waals1893, Cahn1961}.
We note that the physical notion of finite width is particularly relevant for fluids that are near their critical points, or for processes involving moving contact lines, breakup or coalescence, where a sharp interface description will be faced with singularities \citep{Zhang_Mohseni_2018, Eggers1997}.
For this reason, phase field methods are also regarded as \emph{diffuse interface} methods.

@@

In particle-based methods, the central equation to solve is
\begin{equation} 
 \begin{aligned} \label{eq:force-balance}
  {\bm m} \cdot \frac{d{\bm U}}{dt} = {\bm F}^H + {\bm F}^A, 
 \end{aligned}
\end{equation}
where ${\bm m}$ is a generalised mass/moment-of-inertia matrix,
${\bm U}$ is the particle translational/rotational velocity vector,
and ${\bm F}$ represents the force/torque vectors owing to hydrodynamics (denoted with superscript $^H$)
or any applied interactions (denoted with superscript $^A$) that may include contact, electrostatic, van der Waals, or magnetic forces, etc.%
\footnote{In principle, stochastic forces can also be included in the general formulation,
though its numerical treatments demand special care such that the fluctuation-dissipation theorem is satisfied.}
Note that, for $N$ particles in 3D, both ${\bm U}$ and ${\bm F}$ have dimension $6N$, while ${\bm m}$ has size $6N \times 6N$.

@@

There is one more crucial step in Stokesian dynamics.
In the construction of $\mathscr{M}^\infty$, lubrication interactions are not fully accounted for due to the omission of higher-order multipole moments
(hence the superscript $^\infty$).
To correctly include all contributions from the lubrication, $\mathscr{R}$ is calculated as
\begin{equation} \label{eq:sd-lub-correction}
 \begin{aligned}
  \mathscr{R} = (\mathscr{M}^\infty)^{-1} +\mathscr{R}_{2B} - \mathscr{R}_{2B}^\infty,
 \end{aligned}
\end{equation}
where $\mathscr{R}_{2B}^\infty$ denotes the far-field parts of lubrication already embedded in $\mathscr{M}^\infty$,
and the subscript $_{2B}$ indicates that lubrication is mainly a two-body interaction.
See \cite{durlofsky_brady_bossis_1987, Brady_Bossis1988} for more detailed discussions.

Finally, the solution procedure of Stokesian dynamics can be summarized as follows.
\begin{enumerate}
\item Build the grand mobility matrix $\mathscr{M}^\infty$ from the suspension microstructure.
\item Obtain the grand resistance matrix $\mathscr{R}$ by inverting $\mathscr{M}^\infty$ and adding the lubrication correction as in Eq.\ \eqref{eq:sd-lub-correction}.
\item Compute the particle velocity according to Eq.\ \eqref{eq:sd-vel}.
\item Update the particle position using a temporal integration scheme, such as the Euler's method $\bm{x}(t+\Delta t)= \bm{x}(t) +\bm{U} \Delta t$.
\item Go back to Step 1 to continue the time integration unless the desired time step has been reached.
\end{enumerate}

@@

The above expressions for the contact forces and torques can be directly added in the resistance matrices (cf.\ Eq.\ \ref{eq:sd-hydro})
and be solved similarly to the Stokesian dynamics method.
From a computational standpoint, the present method does not require to invert the mobility matrix,
and the grand resistance matrix composed of pair interactions is sparse;
hence, it can be solved much faster numerically.
Alternatively, the particle dynamics can also be obtained by explicitly solving Eq.\ \eqref{eq:force-balance};
namely, first computing all forces and torques to get the acceleration, then integrating in time once to obtain the velocity and twice to update the position.
Similar techniques are sometimes called dissipative particle dynamics or Brownian dynamics, depending on the context \citep{Hoogerbrugge_1992, Groot_Warren_1997}.
Provided that the particle Reynolds number remains much less than unity,
the latter are equivalent to the quasi-static matrix inversion approach used in Stokesian dynamics.

@@

In this chapter, we describe two additional topics that were investigated during the course of the present work.
As side projects, they share with the rest of the thesis in that the flow involved are multiphase and characterised by relatively small length scales; hence, they can be, and have been, studied using similar numerical methods.
Beyond that, there is no further connection.
In the following, we first give the physical motivation of each problem, then summarize the overall understanding and research methodology prior to our study.
The new results are reported in \emph{Paper 5 \& 6}.

@@

Finally, as an outlook, we summarize the lessons learned during all the projects and name a few open questions worth further investigation:
\medskip
